Cooperative Kernels: Safe Multitasking for Blocking Algorithms on GPUs

### 2016-11-10 write-ups/PLDI/paper.tex
Edits.

Figure~\ref{fig:workstealing} shows a simplified version of an OpenCL
kernel that uses a work stealing queue.  \ADComment{Describe some
  aspects of the example, saying that the mutex is not a primitive,
  and can be coded up using the memory model but is not safe due to
  the execution model.  Make it clear that the lock function uses
  busy-waiting.}

\paragraph{Work stealing}
Work stealing is generally used for computational work that can be
splitted into independently processable tasks. The processing of a given
task may create an arbitrary number of new tasks in such a way that it
is typically hard to know the total number of tasks in advance. Work
stealing enables the dynamic balance of task processing over several
processing units, OpenCL workgroups in our case. Each workgroup has an
associated task pool from which it obtains tasks to process and to where
it stores newly created tasks. When a workgroup's pool becomes empty, a
workgroup---rather than terminating---tries to \emph{steal} a task from
an other pool. Hence, workgroups keep on processing tasks until all the
work is done.

@@

HSA provides limited, one-way guarantees,
stating~\cite[p. 46]{HSAprogramming11}: \emph{Work-group A can wait
  for values written by work-group B without deadlock provided ... A
  comes after B in work-group flattened ID order}. However, this
one-way communication is not sufficient so support either the
work-stealing or graph application examples of
Figure~\ref{fig:graphsearch} and \ref{fig:workstealing}. As described
above, both require \emph{symmetric} communication between
threads. \ADComment{Tyler, can you state the HSA guarantee?},
\TSComment{How is this?}~\cite{...}.

@@

For example, many applications using the inter-workgroup barrier
paradigm do not carry workgroup-local information across calls to the
inter-workgroup barrier. Thus the number of workgroups can be adjusted
at every call to the inter-workgroup barrier. Many applications using
the work-stealing paradigm do not carry state between task
computation; thus workgroups can be created or terminated immediately
after ever task computation.

>>>>

\paragraph{Work stealing example}
  \ADComment{Describe some
  aspects of the example, saying that the mutex is not a primitive,
  and can be coded up using the memory model but is not safe due to
  the execution model.  Make it clear that the lock function uses
  busy-waiting.}  
Figure~\ref{fig:workstealing} shows a simplified version of an OpenCL
kernel that uses a work stealing queue.
%
Work stealing is useful when a computation can be split into multiple
tasks that can be processed independently, but where the number of
tasks is not known ahead of time because processing a given task may
create an arbitrary number of new tasks. Work stealing enables dynamic
balancing of tasks across several processing units, OpenCL workgroups
in our case. Each workgroup has an associated task pool from which it
obtains tasks to process, and to which it stores newly created
tasks. When a workgroup's pool becomes empty, a workgroup---rather
than terminating---tries to \emph{steal} a task from an other
pool. Hence, workgroups keep on processing tasks until all the work is
done.

@@

HSA provides limited, one-way guarantees,
stating~\cite[p. 46]{HSAprogramming11}: \emph{Work-group A can wait
  for values written by work-group B without deadlock provided ... A
  comes after B in work-group flattened ID order}. However, this
one-way communication is not sufficient so support either the
work stealing or graph application examples of
Figure~\ref{fig:graphsearch} and \ref{fig:workstealing}. As described
above, both require \emph{symmetric} communication between
threads. \ADComment{Tyler, can you state the HSA guarantee?},
\TSComment{How is this?}~\cite{...}.

@@

For example, many applications using the inter-workgroup barrier
paradigm do not carry workgroup-local information across calls to the
inter-workgroup barrier. Thus the number of workgroups can be adjusted
at every call to the inter-workgroup barrier. Many applications using
the work stealing paradigm do not carry state between task
computation; thus workgroups can be created or terminated immediately
after ever task computation.
